Practical Machine Learning - Cyril Ferdynus
========================================================

Background
-------------------------------

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

Load data
-------------------------------

Data are read from the disk and cleaned (1) Remove cols that are irrelevant for the classification (the first 7 ones)
  (2) Remove cols that contains only NAs or Near Zero Variance

```{r, eval=TRUE}
library(caret)

# Load 
training <- read.csv("E:/Git/PracticalMachineLearning/pml-training.csv", header = TRUE)
test <- read.csv('E:/Git/PracticalMachineLearning/pml-testing.csv')

training$classe <- as.factor(training$classe)

training <- training[, -c(1:7)]
test <- test[, -c(1:7)]

naCount <- apply(training, 2, function(x) {sum(is.na(x))})
nz <- nearZeroVar(training, saveMetrics = TRUE)
training <- training[, which(naCount == 0 & nz$nzv == FALSE)]
```


Splitting data
-------------------------------
Data are splitted in two partitions (60% for the train set and 40% for the test set)

```{r, eval=TRUE}
ti <- createDataPartition(y = training$classe, p=0.6, list=FALSE)
trainPartition <- training[ti, ]
testPartition <- training[-ti, ]

```

Train a machine learning model without Cross Validation
-------------------------------

We chose a random forest algorithm that perform better for this problem

```{r, eval=TRUE}
set.seed(1234)
model <- train(classe ~ ., method = "rf", data = trainPartition)
accuracy <- predict(model , testPartition)
print(confusionMatrix(accuracy, testPartition$classe))
```


Train a machine learning model 10-fold Cross validation
-------------------------------

```{r, eval=TRUE}
set.seed(1234)
tc <- trainControl(method = "repeatedcv", number = 10, repeats = 10)
model_10_CV <- train(classe ~ ., method="rf",  data=trainPartition, trControl = tc)
accuracy_10_CV <- predict(model_10_CV , testPartition)
print(confusionMatrix(accuracy_10_CV, testPartition$classe))
```

Accuracy is just slighlty improved when using Cross-Validation versus not using Cross Validation (99.13% without CV versus 99.15% with CV)
This slight improvement is probably due to that the size of the train sample is very large.
Finally, the two models should missclassify only a few part of the test sample

Predicting 20 test cases
-------------------------------

We predict classes for the 20 test cases using the 10-fold CV algorithm

```{r, eval=TRUE}
pred <- predict(model_10_CV, test)
print(pred)
}
```

Writing results to file for submission
-------------------------------

```{r, eval=TRUE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("E:/Git/PracticalMachineLearning/problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(pred)
```


